### 2015-05-14 whoop.tex
hacks with paragraphs
reduced the font of some figures
merged SMACK unsoundness text

\medskip\noindent\textbf{Concurrent Programming Model }
%
We consider a concurrent programming model where an unbounded number of threads execute a set of pre-defined \emph{thread templates}.  At any given point of execution a certain number of threads are active, each thread executing a particular template.  In the context of device drivers, a thread template corresponds to a driver entry point, and multiple instances of the same thread template may execute concurrently, just as multiple invocations of a single driver entry point may be concurrent.  Further threads may start executing at any point during execution; in the context of device drivers this corresponds to the OS invoking additional driver entry points.\footnote{We do \emph{not} consider the case where one thread spawns another thread, which does not occur in the context of drivers; rather we aim to capture the scenario where additional threads are launched by the environment.}  For ease of presentation only, our model does not feature aggregate data types, pointers or dynamic memory allocation.  These \emph{are} handled by our implementation, and in \S\ref{sec:implementation} we discuss interesting practical issues arising from the handling of a full-blown language.

@@

\begin{figure}
\begin{tabular}{lp{5.5cm}}
\textbf{Statement Form} & \textbf{Notes} \\
\toprule

@@

\medskip\noindent\textbf{Semantics }
%
Let $\ids$ be an infinite set from which dynamic thread ids will be drawn.  The state of a running concurrent program consists of: a valuation of the shared variables, $\sharedvars$; a mapping that associates each mutex in $\mutexes$ with an id from $\ids$, recording which thread currently holds the mutex, or with a special value $\bot \notin \ids$ to indicate that the mutex is not held by any thread; and a list of \emph{threads}.  Each thread consists of an id, drawn from $\ids$, a thread template, $T$, an index indicating the next statement of $T$ to be executed by the thread, and a valuation of the thread private variables, $\variables{T}$.  If multiple threads are instances of the same template $T$, then each thread carries a \emph{separate} valuation of the private variables for this template.

@@

\medskip\noindent\textbf{Lockset Instrumentation }
%
For thread templates $T$ and $U$ (including the possibility that $T$ and $U$ are equal) we wish to check whether it is possible for a thread executing $T$ to race with a thread executing $U$, in the presence of arbitrarily many further concurrently executing threads.

@@

\begin{figure}
\begin{tabular}{ll}
\textbf{Original Statement} & \textbf{Instrumented Statement} \\
\toprule

@@

\medskip\noindent\textbf{Shared State Abstraction }
%
Recall that while our aim is to perform race analysis for pairs of threads, we must be sure to account for possible side-effects due to other threads that are running concurrently.  The instrumentation of Figure~\ref{fig:instrumentation} achieves this via \emph{nondeterminism}: when reading from a shared variable $s$, a nondeterministic value is returned.  This is reflected by the use of a $\havockeyword$ command, which sets its argument to an arbitrary value.  Because all shared state accesses are abstracted in this fashion, it is possible to completely dispense with the shared variables after the lockset instrumentation has been performed.  As a result, when instrumenting a shared variable write, the effect of the write is not explicitly modeled.

\medskip\noindent\textbf{Sequentialisation }
%
The pseudocode of Figure~\ref{fig:sequentialization} shows the sequential program that we analyze in order to prove race-freedom for a pair of thread templates $T$ and $U$. Assuming that $T$ and $U$ have been instrumented using distinct symbols $i$ and $j$, yielding $\instrument{T}{i}$ and $\instrument{U}{j}$, the sequential program operates as follows. First, the read, write and current locksets for $\instrument{T}{i}$ and $\instrument{U}{j}$ are initialized to be empty, and for each shared variable $s$, the locksets $\lockset{s}{i}$ and $\lockset{s}{j}$ are initialized to the full set of mutexes, $\mutexes$.  The main procedures of the instrumented thread templates, $\mainforthread{\instrument{T}{i}}$ and $\mainforthread{\instrument{U}{j}}$, are then executed in turn (the order does not matter, due to renaming). Finally, an assertion checks for consistent use of mutexes: if $s$ is written during execution of $\instrument{T}{i}$ and accessed during execution of $\instrument{U}{j}$, or vice-versa, then the locksets $\lockset{s}{i}$ and $\lockset{s}{j}$ must contain at least one common mutex.

\begin{figure}
\begin{tabular}{l}
$\currentlockset{i} = \emptyset;$ $\hasbeenread{i} = \emptyset;$ $\hasbeenwritten{i} = \emptyset;$ \\
$\currentlockset{j} = \emptyset;$ $\hasbeenread{j} = \emptyset;$ $\hasbeenwritten{j} = \emptyset;$ \\
\texttt{for} $s \in \sharedvars$ \texttt{do} $\lockset{s}{i} = \mutexes;$ $\lockset{s}{j} = \mutexes;$ \medskip
\\

@@

\medskip\noindent\textbf{Soundness }
%
We sketch an argument that if the program of Figure~\ref{fig:sequentialization} is correct, it is impossible for a thread executing template $T$ to race with a thread executing template $U$, under the assumption that the threads are guaranteed to terminate. Let us assume that the program of Figure~\ref{fig:sequentialization} is correct, and suppose (by way of contradiction) that a thread executing $T$ can in fact race with a thread executing $U$, on some shared variable $s$.  By our hypothesis that the program of Figure~\ref{fig:sequentialization} is correct, and that the threads terminate, the assertion checked at the end of the program guarantees at least one mutex, say $m$, belongs to both $\lockset{s}{i}$ and $\lockset{s}{j}$.  By the definition of a lockset (and according to the manner in which shared accesses are instrumented in Figure~\ref{fig:instrumentation}), this means that $m$ is held during every access to $s$ by both $\instrument{T}{i}$ and $\instrument{U}{j}$. As a result, $m$ must be unlocked and locked between the two accesses, which contradicts that the pair of accesses is racing.

@@

\medskip\noindent\textbf{Architecture }
%
Figure~\ref{fig:whoop} depicts the \whoop toolchain. Initially, \whoop accepts a Linux driver written in C, together with an environmental model\footnote{This consists of stub header files modeling relevant Linux kernel APIs.} that is required to ``close'' the driver and allow it to be subsequently analyzed for data races. Both the driver and the Linux model are passed to the translation phase of \whoop (see Figure~\ref{fig:whoop} -- A), which uses three LLVM\footnote{\url{http://llvm.org}}-based tools, Chauffeur\footnote{\url{https://github.com/mc-imperial/chauffeur}}, Clang\footnote{\url{http://clang.llvm.org}} and SMACK\footnote{\url{https://github.com/smackers/smack}}~\cite{rakamaric2014smack}, to translate the original driver into an abstract program written in the Boogie intermediate verification language (IVL)~\cite{deline2005boogiepl}, a simple imperative language with well-defined semantics that is used as the input to a number of cutting-edge verifiers (e.g.\ Boogie and \corral).

@@

\medskip\noindent\textbf{Extracting Entry Point Information }
%
Chauffeur is a Clang-based frontend that traverses the abstract syntax tree (AST) of the original driver (using a Clang visitor pass) and extracts all entry point identifier names, together with the identifier names of their corresponding kernel API functions. Linux drivers define entry points in a standard way (see Figure~\ref{fig:entrypoints} for an example of how the generic\_nvram driver defines the entry points for the \texttt{file\_operations} API); Chauffeur identifies these definitions in the AST and outputs the relevant information in an XML file, which is subsequently parsed by \whoop and used during the instrumentation.

@@

\medskip\noindent\textbf{Translation for Verification }
%
Next, the driver is compiled by Clang to LLVM-IR~\cite{lattner2004llvm}, a low-level assembly-like language in single static assignment (SSA) form. Function calls (e.g.\ for locking and unlocking) are preserved in this translation, which is useful for our lockset instrumentation.

@@

To achieve scalability, SMACK uses a \emph{split} memory model that exploits an alias analysis to soundly partition memory locations into non-overlapping equivalence classes that do not alias. This has been shown to lead to more tractable verification compared with a \emph{monolithic} model where the heap is considered to be an array of bytes~\cite{rakamaric2009scalable}. The split memory model is based on \emph{memory regions}, which are maps of integers that model the heap. A benefit of using this model is that distinct memory regions denote disjoint sections of the heap. We leverage this knowledge inside \whoop to guide and optimize our lockset instrumentation and analysis, and to create a fine-grained context-switch instrumentation as discussed in \S\ref{corral}. Note that the SMACK's memory model can potentially be unsound in (typically rare) situations where programs use unaligned byte-level memory accesses. The latest release of SMACK addresses this issue, and as future work we plan to integrate the latest SMACK with \whoop to further improve soundness.

\medskip\noindent\textbf{Identifying Locks }
%
When the instrumentation phase begins, \whoop performs an inter-procedural static analysis (on the Boogie IVL source code of each entry point) to identify all available locks and rewrite each one (both at declaration and at all access sites) to a unique constant Boogie variable. The reason behind this transformation, is that representing all locks statically, instead of their original SMACK pointer-based representation, helps \whoop perform various internal instrumentation and optimization passes.
%
Currently, \whoop only supports mutexes and spinlocks that are available in the Linux kernel APIs. However, it is relatively easy to enhance our tool with knowledge of other locking primitives. If \whoop cannot infer a lock, e.g. because it was created dynamically or was indexed from an array of locks, it will exit with a warning. The are two reasons behind this: (i) it is arguably hard to detect such locks using static analysis; and (ii) a small, fixed number of locks is advocated by Linux experts as good practice when developing drivers~\cite{corbet2005linux}.

\medskip\noindent\textbf{Watchdog Race-Checking Instrumentation }
%
Data race detection is performed by introducing sets containing the locks that are consistently held by each shared variable and sets containing all shared variables that are read and written (see \S\ref{sec:symbolicpairwise} and the instrumentation of Figure~\ref{fig:instrumentation}). These sets can be modeled directly in Boogie as characteristic functions, using maps. However, this requires the use of quantifiers to express properties related to set contents.  For instance, to express that a set $X$ of elements of type $A$ is empty, where $X$ is represented as a map from $A$ to $\mathsf{Bool}$, we would require the quantified expression $\forall a : A : \neg X[a]$.  It is well known that automated theorem proving in the presence of quantified constraints is challenging, and that theorem provers such as Z3~\cite{de2008z3} are often much more effective when quantifiers are avoided.

@@

\medskip\noindent\textbf{Generating Loop and Procedure Summaries }
%
Early in the development of \whoop we experimented with reasoning about recursion-free drivers by resorting to full inlining.  We found that this did not scale to large drivers: when we tried to apply \whoop to the r8169 ethernet driver (see \S\ref{evaluation}), we quickly exhausted the memory limits of our experimental platform as the driver exhibited recursion, which cannot be fully inlined and blows up.

@@

\medskip\noindent\textbf{Verification and Error Reporting }
%
For each pair of entry points, the instrumented sequential program, equipped with procedure and loop summaries, is sent to the Boogie verification engine. Boogie generates a VC, for each procedure in the program, and discharges it to the Z3~\cite{de2008z3} theorem prover.  In particular, the verification for the root-level procedure, encoding the sequential program sketched in Figure~\ref{fig:sequentialization}, encodes the race-freedom check for the entry point pair. Successful verification implies that the entry point pair is free of data races, while an error (i.e.\ counterexample) denotes a \emph{potential} data race and is reported to the user. To improve usability, \whoop has a built-in error reporter that matches counterexamples to source code. The following is a race that \whoop found and reported for the example of Figure~\ref{fig:data_race_example}:

@@

\medskip\noindent\textbf{Optimizations }
%
We have implemented various optimizations to increase the precision and performance of \whoop.  We comment on the two most effective optimizations.

@@

\medskip\noindent\textbf{Practical Assumptions Related to Soundness}
%
\whoop is ``soundy''\footnote{\url{http://soundiness.org/}}~\cite{soundiness}: it aims in principle to perform a sound analysis that can prove absence of races, but suffers from some known sources of unsoundness, which we now comment on.

@@

As well as inheriting soundness issues arising from currently unknown bugs in \whoop and in the external tools that \whoop relies on, we acknowledge that SMACK is subject to sources of unsoundness (e.g.\ SMACK models integers as an infinite set, rather than as a finite set of bit-vectors) and that the combination of Clang and SMACK commits our approach to specific choices related to undefined and implementation-defined aspects of the C language when translating to Boogie.  However, \whoop makes no fundamental assumptions related to these translation decisions, meaning that a more accurate C-to-Boogie translation would automatically lead to a more accurate analysis with \whoop.

\medskip\noindent\textbf{Limitations }
%
As a lockset analyzer, \whoop can be imprecise because a violation of the locking discipline does not always correspond to a real race (e.g.\ when lockfree synchronization is used). \whoop also over-approximates drivers, which can be another source of imprecision. Furthermore, the tool does not check for dynamically created locks or for locks provided by external libraries, although the later could be addressed by providing a mechanism for users to declare custom locks. We also do not currently treat interrupt handlers in a special way; we just assume that they can execute concurrently with any entry point. One way to address this is to model interrupt-specific kernel functions (e.g. for enabling/disabling interrupts).

>>>>

\noindent\textbf{Concurrent Programming Model }
%
We consider a concurrent programming model where an unbounded number of threads execute a set of pre-defined \emph{thread templates}.  At any given point of execution a certain number of threads are active, each thread executing a particular template.  In the context of device drivers, a thread template corresponds to a driver entry point, and multiple instances of the same thread template may execute concurrently, just as multiple invocations of a single driver entry point may be concurrent.  Further threads may start executing at any point during execution; in the context of device drivers this corresponds to the OS invoking additional driver entry points.\footnote{We do \emph{not} consider the case where one thread spawns another thread, which does not occur in the context of drivers; rather we aim to capture the scenario where additional threads are launched by the environment.}  For ease of presentation only, our model does not feature aggregate data types, pointers or dynamic memory allocation.  These \emph{are} handled by our implementation, and in \S\ref{sec:implementation} we discuss interesting practical issues arising from the handling of a full-blown language.

@@

\begin{figure}
\footnotesize
\center
\begin{tabular}{lp{5.5cm}}
\textbf{Statement Form} & \textbf{Notes} \\
\toprule

@@

\noindent\textbf{Semantics }
%
Let $\ids$ be an infinite set from which dynamic thread ids will be drawn.  The state of a running concurrent program consists of: a valuation of the shared variables, $\sharedvars$; a mapping that associates each mutex in $\mutexes$ with an id from $\ids$, recording which thread currently holds the mutex, or with a special value $\bot \notin \ids$ to indicate that the mutex is not held by any thread; and a list of \emph{threads}.  Each thread consists of an id, drawn from $\ids$, a thread template, $T$, an index indicating the next statement of $T$ to be executed by the thread, and a valuation of the thread private variables, $\variables{T}$.  If multiple threads are instances of the same template $T$, then each thread carries a \emph{separate} valuation of the private variables for this template.

@@

\noindent\textbf{Lockset Instrumentation }
%
For thread templates $T$ and $U$ (including the possibility that $T$ and $U$ are equal) we wish to check whether it is possible for a thread executing $T$ to race with a thread executing $U$, in the presence of arbitrarily many further concurrently executing threads.

@@

\begin{figure}
\footnotesize
\center
\begin{tabular}{ll}
\textbf{Original Statement} & \textbf{Instrumented Statement} \\
\toprule

@@

\noindent\textbf{Shared State Abstraction }
%
Recall that while our aim is to perform race analysis for pairs of threads, we must be sure to account for possible side-effects due to other threads that are running concurrently.  The instrumentation of Figure~\ref{fig:instrumentation} achieves this via \emph{nondeterminism}: when reading from a shared variable $s$, a nondeterministic value is returned.  This is reflected by the use of a $\havockeyword$ command, which sets its argument to an arbitrary value.  Because all shared state accesses are abstracted in this fashion, it is possible to completely dispense with the shared variables after the lockset instrumentation has been performed.  As a result, when instrumenting a shared variable write, the effect of the write is not explicitly modeled.

\noindent\textbf{Sequentialisation }
%
The pseudocode of Figure~\ref{fig:sequentialization} shows the sequential program that we analyze in order to prove race-freedom for a pair of thread templates $T$ and $U$. Assuming that $T$ and $U$ have been instrumented using distinct symbols $i$ and $j$, yielding $\instrument{T}{i}$ and $\instrument{U}{j}$, the sequential program operates as follows. First, the read, write and current locksets for $\instrument{T}{i}$ and $\instrument{U}{j}$ are initialized to be empty, and for each shared variable $s$, the locksets $\lockset{s}{i}$ and $\lockset{s}{j}$ are initialized to the full set of mutexes, $\mutexes$.  The main procedures of the instrumented thread templates, $\mainforthread{\instrument{T}{i}}$ and $\mainforthread{\instrument{U}{j}}$, are then executed in turn (the order does not matter, due to renaming). Finally, an assertion checks for consistent use of mutexes: if $s$ is written during execution of $\instrument{T}{i}$ and accessed during execution of $\instrument{U}{j}$, or vice-versa, then the locksets $\lockset{s}{i}$ and $\lockset{s}{j}$ must contain at least one common mutex.

\begin{figure}
\footnotesize
\begin{tabular}{l}
$\currentlockset{i} = \emptyset;$ $\hasbeenread{i} = \emptyset;$ $\hasbeenwritten{i} = \emptyset;$ \\
$\currentlockset{j} = \emptyset;$ $\hasbeenread{j} = \emptyset;$ $\hasbeenwritten{j} = \emptyset;$ \\
\texttt{for} $s \in \sharedvars$ \texttt{do} $\lockset{s}{i} = \mutexes;$ $\lockset{s}{j} = \mutexes;$ \medskip
\\

@@

\noindent\textbf{Soundness }
%
We sketch an argument that if the program of Figure~\ref{fig:sequentialization} is correct, it is impossible for a thread executing template $T$ to race with a thread executing template $U$, under the assumption that the threads are guaranteed to terminate. Let us assume that the program of Figure~\ref{fig:sequentialization} is correct, and suppose (by way of contradiction) that a thread executing $T$ can in fact race with a thread executing $U$, on some shared variable $s$.  By our hypothesis that the program of Figure~\ref{fig:sequentialization} is correct, and that the threads terminate, the assertion checked at the end of the program guarantees at least one mutex, say $m$, belongs to both $\lockset{s}{i}$ and $\lockset{s}{j}$.  By the definition of a lockset (and according to the manner in which shared accesses are instrumented in Figure~\ref{fig:instrumentation}), this means that $m$ is held during every access to $s$ by both $\instrument{T}{i}$ and $\instrument{U}{j}$. As a result, $m$ must be unlocked and locked between the two accesses, which contradicts that the pair of accesses is racing.

@@

\noindent\textbf{Architecture }
%
Figure~\ref{fig:whoop} depicts the \whoop toolchain. Initially, \whoop accepts a Linux driver written in C, together with an environmental model\footnote{This consists of stub header files modeling relevant Linux kernel APIs.} that is required to ``close'' the driver and allow it to be subsequently analyzed for data races. Both the driver and the Linux model are passed to the translation phase of \whoop (see Figure~\ref{fig:whoop} -- A), which uses three LLVM\footnote{\url{http://llvm.org}}-based tools, Chauffeur\footnote{\url{https://github.com/mc-imperial/chauffeur}}, Clang\footnote{\url{http://clang.llvm.org}} and SMACK\footnote{\url{https://github.com/smackers/smack}}~\cite{rakamaric2014smack}, to translate the original driver into an abstract program written in the Boogie intermediate verification language (IVL)~\cite{deline2005boogiepl}, a simple imperative language with well-defined semantics that is used as the input to a number of cutting-edge verifiers (e.g.\ Boogie and \corral).

@@

\noindent\textbf{Extracting Entry Point Information }
%
Chauffeur is a Clang-based frontend that traverses the abstract syntax tree (AST) of the original driver (using a Clang visitor pass) and extracts all entry point identifier names, together with the identifier names of their corresponding kernel API functions. Linux drivers define entry points in a standard way (see Figure~\ref{fig:entrypoints} for an example of how the generic\_nvram driver defines the entry points for the \texttt{file\_operations} API); Chauffeur identifies these definitions in the AST and outputs the relevant information in an XML file, which is subsequently parsed by \whoop and used during the instrumentation.

@@

\noindent\textbf{Translation for Verification }
%
Next, the driver is compiled by Clang to LLVM-IR~\cite{lattner2004llvm}, a low-level assembly-like language in single static assignment (SSA) form. Function calls (e.g.\ for locking and unlocking) are preserved in this translation, which is useful for our lockset instrumentation.

@@

To achieve scalability, SMACK uses a \emph{split} memory model that exploits an alias analysis to soundly partition memory locations into non-overlapping equivalence classes that do not alias. This has been shown to lead to more tractable verification compared with a \emph{monolithic} model where the heap is considered to be an array of bytes~\cite{rakamaric2009scalable}. The split memory model is based on \emph{memory regions}, which are maps of integers that model the heap. A benefit of using this model is that distinct memory regions denote disjoint sections of the heap. We leverage this knowledge inside \whoop to guide and optimize our lockset instrumentation and analysis, and to create a fine-grained context-switch instrumentation as discussed in \S\ref{corral}.

\noindent\textbf{Identifying Locks }
%
When the instrumentation phase begins, \whoop performs an inter-procedural static analysis (on the Boogie IVL source code of each entry point) to identify all available locks and rewrite each one (both at declaration and at all access sites) to a unique constant Boogie variable. The reason behind this transformation, is that representing all locks statically, instead of their original SMACK pointer-based representation, helps \whoop perform various internal instrumentation and optimization passes.
%
Currently, \whoop only supports mutexes and spinlocks that are available in the Linux kernel APIs. However, it is relatively easy to enhance our tool with knowledge of other locking primitives. If \whoop cannot infer a lock, e.g. because it was created dynamically or was indexed from an array of locks, it will exit with a warning. The are two reasons behind this: (i) it is arguably hard to detect such locks using static analysis; and (ii) a small, fixed number of locks is advocated by Linux experts as good practice when developing drivers~\cite{corbet2005linux}.

\noindent\textbf{Watchdog Race-Checking Instrumentation }
%
Data race detection is performed by introducing sets containing the locks that are consistently held by each shared variable and sets containing all shared variables that are read and written (see \S\ref{sec:symbolicpairwise} and the instrumentation of Figure~\ref{fig:instrumentation}). These sets can be modeled directly in Boogie as characteristic functions, using maps. However, this requires the use of quantifiers to express properties related to set contents.  For instance, to express that a set $X$ of elements of type $A$ is empty, where $X$ is represented as a map from $A$ to $\mathsf{Bool}$, we would require the quantified expression $\forall a : A : \neg X[a]$.  It is well known that automated theorem proving in the presence of quantified constraints is challenging, and that theorem provers such as Z3~\cite{de2008z3} are often much more effective when quantifiers are avoided.

@@

\noindent\textbf{Generating Loop and Procedure Summaries }
%
Early in the development of \whoop we experimented with reasoning about recursion-free drivers by resorting to full inlining.  We found that this did not scale to large drivers: when we tried to apply \whoop to the r8169 ethernet driver (see \S\ref{evaluation}), we quickly exhausted the memory limits of our experimental platform as the driver exhibited recursion, which cannot be fully inlined and blows up.

@@

\noindent\textbf{Verification and Error Reporting }
%
For each pair of entry points, the instrumented sequential program, equipped with procedure and loop summaries, is sent to the Boogie verification engine. Boogie generates a VC, for each procedure in the program, and discharges it to the Z3~\cite{de2008z3} theorem prover.  In particular, the verification for the root-level procedure, encoding the sequential program sketched in Figure~\ref{fig:sequentialization}, encodes the race-freedom check for the entry point pair. Successful verification implies that the entry point pair is free of data races, while an error (i.e.\ counterexample) denotes a \emph{potential} data race and is reported to the user. To improve usability, \whoop has a built-in error reporter that matches counterexamples to source code. The following is a race that \whoop found and reported for the example of Figure~\ref{fig:data_race_example}:

@@

\noindent\textbf{Optimizations }
%
We have implemented various optimizations to increase the precision and performance of \whoop.  We comment on the two most effective optimizations.

@@

\noindent\textbf{Practical Assumptions Related to Soundness}
%
\whoop is ``soundy''\footnote{\url{http://soundiness.org/}}~\cite{soundiness}: it aims in principle to perform a sound analysis that can prove absence of races, but suffers from some known sources of unsoundness, which we now comment on.

@@

As well as inheriting soundness issues arising from currently unknown bugs in \whoop and in the external tools that \whoop relies on, we acknowledge that: (i) SMACK is subject to sources of unsoundness, e.g.\ it models integers as an infinite set (rather than as a finite set of bit-vectors), and its memory model can potentially be unsound in (typically rare) situations where programs use unaligned byte-level memory accesses; and (ii) that the combination of Clang and SMACK commits our approach to specific choices related to undefined and implementation-defined aspects of the C language when translating to Boogie.  However, \whoop makes no fundamental assumptions related to these translation decisions, meaning that a more accurate C-to-Boogie translation would automatically lead to a more accurate analysis with \whoop.

\noindent\textbf{Limitations }
%
As a lockset analyzer, \whoop can be imprecise because a violation of the locking discipline does not always correspond to a real race (e.g.\ when lockfree synchronization is used). \whoop also over-approximates drivers, which can be another source of imprecision. Furthermore, the tool does not check for dynamically created locks or for locks provided by external libraries, although the later could be addressed by providing a mechanism for users to declare custom locks. We also do not currently treat interrupt handlers in a special way; we just assume that they can execute concurrently with any entry point. One way to address this is to model interrupt-specific kernel functions (e.g. for enabling/disabling interrupts).
